### 2019-11-14 thesis.tex
kinetic theory

In the context of plasma kinetic theory, processes involving waves
and kinetic instabilities, the so-called \emph{collective processes},
are almost exclusively associated with situations in which the effects
of collisional dissipation can be neglected, what happens when the
studied phenomenon evolves in a much shorter time-scale than the
collisional relaxation time of the plasma in question \cite{akhi1967}.
In such cases, the plasma dynamics may be described by the Vlasov-Maxwell
system, a complex set of coupled equations whose
%which 
% LFZ190822: Posso estar enganado, mas me parece que aqui seria mesmo
% "whose", como estava antes, e não "which".
% Sabrina: resolvi colocar "whose" de novo depois de uma rápida, mas
% controversa, pesquisa no google
solution invariably will
depend on some degree of approximation. In the presence of small
amplitude oscillations in a fully ionized plasma, we can make use
of perturbation theory in order to solve such complicated system
\cite{klimo}. From this method, one may obtain simpler approximations
like the linear theory and the quasilinear formalism, and also
more complex formulations, like the weak turbulence theory, which
takes into account low-order nonlinearities on the plasma dynamics.

@@

In the quasilinear theory, low-order nonlinear terms are kept in
the equation for the time evolution of the velocity distribution
function. The velocity distribution will then evolve in time through
a process of diffusion in velocity space, in which the diffusion
coefficient depends on the spectral energy density of the waves.
The dispersion relation is formally given by the same expression
obtained under the linear theory, except that now it depends on a
time-dependent velocity distribution function, which must vary in a
much slower time-scale than the period of the plasma oscillations.
Under such condition, the gradual changing on the velocity space
shape slowly modifies the shape of the wave spectrum, whose
% LFZ190822: Não seria mesmo "whose"?
% Sabrina: ok
new form will alter the velocity distribution function and so
on, until a new quasi-stationary equilibrium state is attained
\cite{gurnett2017}. However, depending on the phenomenon being
described this saturation state reached under the quasilinear
approximation might keep on evolving if higher order nonlinearities are taken into account
\cite{akhi2}. This kind of nonlinear analysis is given by the
\emph{weak turbulence theory}, which is the next step in this chain
of perturbative approximations.

@@

\chapter{Kinetic theory of plasmas}
\label{cha:kin-theo}
Plasmas are ionized gases that exhibit collective behavior. Usually composed
of positive and negative charges and neutral particles, a plasma is said to
be \emph{neutral} when the quantity of particles with opposite charges is the
same~\cite{klimo}, and \emph{nonneutral} when it is constituted - or have a
significant excess - of one kind of charge~\cite{DavidsonNonNeutral}. In this
work, it will be considered a fully ionized hydrogen plasma, which is overall
neutral. In such case, the density of protons and electrons per volume unity
is the same, therefore $n_e=n_i=n$.

Electrons and ions are sources of electric and magnetic fields. Due to the
long-range nature of these fields, particles can interact simultaneously, at
distance, with several other particles in the system. The effective range of
the interactions among charges is determined by the \emph{Debye length}
\begin{equation}
  \label{debye}
  \lambda_{De}=\sqrt{\frac{k_BT_e}{4\pi n e^2}},
\end{equation}
where $T_e$ is the electron temperature, $k_B$ the Boltzmann constant, $n$
the plasma density and $e$ the charge. This means that a specific charge
will interact effectively only with other charges that are inside of a
virtual sphere with radius given by $\lambda_{De}$, being almost completely
shielded from the influence of particles outside the \emph{Debye sphere}.

The kinetic theory is the most fundamental description of the plasma
dynamics. Based on the concepts and methods of statistical mechanics,
together with the Maxwell equations, it provides a formal self-consistent
evolution of the plasma processes. This formalism takes into account the
variations of the velocity distribution function $f({\bf r},{\bf v},t)$
in the phase-space for each particle species. Statistically, this means
that the product between the distribution function and the volume element
in the hexa-dimensional phase-space $f_\alpha({\bf r},{\bf v},t)d^3r d^3v$
represents the probability of finding a particle of species
$\alpha$ in a volume element $d^3r$ around ${\bf r}$ and in a
velocity volume element $d^3v$ around ${\bf v}$, at the instant
$t$ \cite{klimo,klimon}.

@@

\chapter{Weak turbulence theory}
\label{cha:weak-turb}
The weak turbulence theory is what we may call a ``borderline approach''
because, despite the fact that it deals with some more intense
instabilities, we must assure that these fluctuations are still inside
the weak turbulence limit, i.e., if the turbulence is weak enough to be
treated by a perturbative method. Such boundary is defined by a comparison
between the averaged kinetic energy of the particles per volume unity
$\calE_{kin}=\bar n m<v^2>/2$, and the energy density $\calE_f$, which
is associated with the fluctuations of the perturbative electric field.
Thus, by ``weak turbulence'' we mean
\begin{equation}
  \label{eq:weak}
  \calE_f \ll \calE_{kin},
\end{equation}
that is,
% LFZ180216: Não fica melhor com vírgula depois de "that is"?
% Sabrina: OK
the energy density of the fluctuations must be much smaller
% LFZ180216: Ou simplesmente "much smaller"?
% Sabrina: OK
than the kinetic energy density of the particles. This condition is
related to the slow wave-growing assumption \cite{david}, which is
central for the further development of nonlinear kinetic equations,
as will be discussed further.

@@

\chapter{Microscopic equations for a fully ionized plasma}
\label{appF}
Plasmas are ionized gases that exhibit collective behavior. Usually composed
of positive and negative charges and neutral particles, a plasma is said to
be \emph{neutral} when the quantity of particles with opposite charges is the
same~\cite{klimo}, and \emph{nonneutral} when it is constituted - or have a
significant excess - of one kind of charge~\cite{DavidsonNonNeutral}. In this
work, it will be considered a fully ionized hydrogen plasma, which is overall
neutral, with the density of protons and electrons being the same, therefore
$n_e=n_i=n$.

Electrons and ions are sources of electric and magnetic fields. Due to the
long-range nature of these fields, particles can interact simultaneously, at
distance, with several other particles in the system. The effective range of
the interactions among charges is determined by the \emph{Debye length}
\begin{equation}
  \label{debye}
  \lambda_{De}=\sqrt{\frac{k_BT_e}{4\pi n e^2}},
\end{equation}
where $T_e$ is the electron temperature, $k_B$ the Boltzmann constant, $n$
the plasma density and $e$ the charge. This means that a specific charge
will interact effectively only with other charges that are inside of a
virtual sphere with radius given by $\lambda_{De}$, being almost completely
shielded from the influence of particles outside the \emph{Debye sphere}.

>>>>

In the context of plasma kinetic theory, processes involving waves
and kinetic instabilities, the so-called \emph{collective processes},
are almost exclusively associated with situations in which the effects
of collisional dissipation can be neglected, what happens when the
studied phenomenon evolves in a much shorter time-scale than the
collisional relaxation time of the plasma in question \cite{akhi1967}.
In such cases, the plasma dynamics may be described by the Vlasov-Maxwell
system, a complex set of coupled equations whose solution invariably
will depend on some degree of approximation. In the presence of small
amplitude oscillations in a fully ionized plasma, we can make use
of perturbation theory in order to solve such complicated system
\cite{klimo}. From this method, one may obtain simpler approximations
like the linear theory and the quasilinear formalism, and also
more complex formulations, like the weak turbulence theory, which
takes into account low-order nonlinearities on the plasma dynamics.

@@

In the quasilinear theory, low-order nonlinear terms are kept in
the equation for the time evolution of the velocity distribution
function. The velocity distribution will then evolve in time through
a process of diffusion in velocity space, in which the diffusion
coefficient depends on the spectral energy density of the waves.
The dispersion relation is formally given by the same expression
obtained under the linear theory, except that now it depends on a
time-dependent velocity distribution function, which must vary in a
much slower time-scale than the period of the plasma oscillations.
Under such condition, the gradual changing on the velocity space
shape slowly modifies the shape of the wave spectrum, whose new form
will alter the velocity distribution function and so on, until a new
quasi-stationary equilibrium state is attained \cite{gurnett2017}.
However, depending on the phenomenon being described this saturation
state reached under the quasilinear approximation might keep on evolving
if higher order nonlinearities are taken into account \cite{akhi2}.
This kind of nonlinear analysis is given by the
\emph{weak turbulence theory}, which is the next step in this chain
of perturbative approximations.

@@

\chapter{Kinetic theory of plasmas}
\label{cha:kin-theo}
Plasmas are ionized gases that exhibit collective behavior. Composed of
positive and negative charges and neutral particles, a plasma is said to
be \emph{neutral} when the quantity of particles with opposite charges is
the same~\cite{klimon}, and \emph{nonneutral} when it is constituted - or
have a significant excess - of one kind of charge~\cite{DavidsonNonNeutral}.
For the present work, we assume a fully ionized hydrogen plasma, which is
overall neutral and does not contain any neutral particle. In such a case,
half the total number of particles is given by protons and the other half
by electrons.

Electrons and ions are sources of electric and magnetic fields. Due to the
long-range nature of these fields, particles can interact simultaneously, at
distance, with several other particles in the system. The effective range of
these interactions is determined by the \emph{Debye length}
\begin{equation}
  \label{debye}
  \lambda_{De}=\sqrt{\frac{k_BT_e}{4\pi n e^2}},
\end{equation}
where $T_e$ is the electron temperature, $k_B$ the Boltzmann constant, $n$
the plasma density and $e$ the electron charge. This means that a specific
charge will interact effectively only with other charges that are inside of a
virtual sphere with radius given by $\lambda_{De}$, being almost completely
shielded from the influence of particles outside the \emph{Debye sphere}.
The self-consistent interaction between charged particles and electric and
magnetic fields, creates fluctuations in the local neutrality of the plasma,
giving rise to the series of oscillations and complex wave phenomena that
characterize the plasma’s collective behavior~\cite{chen}.

Describing the precise state of a plasma, in a given time, would require
the knowledge of the $6N$ phase-space coordinates of each particle, and also
the microscopic amplitudes of the electric and magnetic fields at each point
in space. Such exact representation is clearly a non-practical approach for
a many-body system like the plasma. However, this microscopic formulation
can be approximated to a macroscopic description by taking the average of
the microscopic quantities and making suitable assumptions to truncate the
correlation functions at an appropriate point~\cite{klimon}. This will lead
to the plasma kinetic equations, which is the first of a hierarchy of
approximations that also includes the multi-fluid and magnetohydrodynamic
theories. The application of each one of these formulations will depend on
the context and the scale of the phenomena being described~\cite{chen}.
In this work we are interested in small-scale processes that depend on
velocity-space properties, requiring a kinetic description of the plasma.

The kinetic theory is the most fundamental description of the plasma
dynamics. Based on the concepts and methods of statistical mechanics,
together with the Maxwell equations, it provides a formal self-consistent
evolution of the plasma processes. This formalism takes into account the
variations of the velocity distribution function $f({\bf r},{\bf v},t)$
in the phase-space for each particle species. Statistically, this means
that the product between the distribution function and the volume element
in the hexa-dimensional phase-space $f_\alpha({\bf r},{\bf v},t)d^3r d^3v$
represents the probability of finding a particle of species $\alpha$ in a
volume element $d^3r$ around ${\bf r}$ and in a velocity volume element
$d^3v$ around ${\bf v}$, at the instant $t$ \cite{klimon}.

In a first approximation, the plasma dynamics is described by the
Boltzmann equation
\begin{equation}
\frac{\partial f_\alpha({\bf r},{\bf v},t)}{\partial t}
+{\bf v \cdot \nabla_r}f_\alpha({\bf r},{\bf v},t)
 +\frac{{\bf F}}{m_\alpha} {\bf\cdot \nabla_v}f_\alpha({\bf r},{\bf v},t)
=\left(\frac{\partial f_\alpha ({\bf r},{\bf v},t)}{\partial t}\right)_{coll}.
\label{boltz}
\end{equation}

@@

\chapter{Weak turbulence theory}
\label{cha:weak-turb}
The weak turbulence theory is what we may call a ``borderline approach''
because, despite the fact that it deals with some more intense
instabilities, we must assure that these fluctuations are still inside
the weak turbulence limit, i.e., if the turbulence is weak enough to be
treated by a perturbative method. Such boundary is defined by a comparison
between the averaged kinetic energy of the particles per volume unity
${\cal E}_{kin}=\bar n m<v^2>/2$, and the energy density ${\cal E}_f$, which
is associated with the fluctuations of the perturbative electric field.
Thus, by ``weak turbulence'' we mean
\begin{equation}
  \label{eq:weak}
  {\cal E}_f \ll {\cal E}_{kin},
\end{equation}
that is,
% LFZ180216: Não fica melhor com vírgula depois de "that is"?
% Sabrina: OK
the energy density of the fluctuations must be much smaller
% LFZ180216: Ou simplesmente "much smaller"?
% Sabrina: OK
than the kinetic energy density of the particles. This condition is
related to the slow wave-growing assumption \cite{david}, which is
central for the further development of nonlinear kinetic equations,
as will be discussed further.

@@
